### 2022-05-29 article.tex
All refs are in

As an example, we'll use a pseudorandomly selected paper
from [Ahsan et al., 2017]. The data used by that paper is
from a population study called the Northern Swedish Population
Health Study (NSPHS) that started in 2010 [Igl et al., 2010]. 
The approximately 1000 participants were initially mostly surveyed
about lifestyle [Igl et al., 2010] and follow-up studies
provided the type of data relevant for this paper, 
which are (1) the genetic data [Johansson et al., 2013],
(2) the concentration of certain proteins in the 
blood [Enroth et al., 2014][Enroth et al., 2015].

@@

The first type of data, the (final form of the) genetic data,
consists out of single nucleotide polymorphisms (SNPs).
SNPs consist out of a name, a position and a nucleotide. 
DNA (organized into chromosomes and present in every (nucleated) cell, 
see figure F3) consists out of many nucleotides, of which there are four types,
called adenosine, cytosine, guanine and thyrosine, all commonly abreviated
as A, C, G and T respectively.
One SNP example is, that 'rs12133641' is a SNP located at position 154,428,283,
where some people have a certain nucleotide. In this case, 
67 percent of the people within this study have an A,
and 33 percent have a G (also from [Ahsan et al., 2017], Table S3).

@@

The field of genetic epidemiology looks -among others- for
correlations between genetic data and biological traits.
For example, figure F10 (from [Ahsan et al., 2017]) shows that
SNP rs12133641 is highly correlated (p value is 3.0 * 10e-73, 
961 individuals) with protein IL6RA. What this results does
not yet teach us, is how this correlation works, yet
figure F20 shows us the direction of the association:
the X axis shows the possible genetic make-ups (aka the genotype) of the individuals,
where 0 denotes the individuals with genotype
AA (the individual inherited one adenosine 
from his/here mother and one adenosine from his/her father), 
1.0 denotes AG (one A is inherited from one parent, 
where the G is inherited from the other parent) and 2.0 denotes GG.
From Figure F20 we can conclude that, on average,
the more guanines are inherited at that SNPs location,
the higher concentration of IL6RA can be found in a human's blood.

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{ahsan2017relative_table_2_sub.png}
  \caption{
    Figure F10 shows an example result of a genetic epideological research.
    It shows that the SNP named rs12133641 (located at position 154,428,283
    of chromosome 1) is highly correlated (p value is 3.0 * 10e-73, 
    961 individuals) to the concentration of the protein IL6RA, as measured
    in blood. The table is a simplified result from [Ahsan et al., 2017].
  }
  \label{fig:ahsan2017relative_table_2_sub}
\end{figure}

@@

To quote a review on reproducibility [Peng and Hicks, 2021], 
'Reproducibility is typically thwarted by a 
lack of availability of the original data and computer code'.
In the case of genetic epidemiology, it is a given that the original
data cannot be published as this data is sensitive.

@@

The line of code is simplified (as the result of the T test is never stored),
but the code is correct: 
`t.test` is indeed the name of an R function to do a T-test
and it is reasonable to assume that this code, when found in the code
of an article, is assumed to be correct.

Here, however, we see a mismatch between the English description and the code:
by default, the R function `t.test` does a **two**-tailed T-test.
A consequence of this fictional example is that the published p-values are
higher than needed, resulting in less significant findings, which
results in less conclusions drawn from a paper.

@@

The code that does the work in a computational experiment should
always be published together with a scholarly article.
For research to be reproducible, one ideally has access to
both the data used and the code.
In some fields, such as genetic epidemiology, the data is
sensitive, hence cannot be released,
yet there are methods being devised to run code on sensitive
data with assured privacy [Zhang et al., 2016][Azencott, 2018].

>>>>

As an example, we'll use a pseudorandomly selected paper
from \cite{ahsan2017relative}. The data used by that paper is
from a population study called the Northern Swedish Population
Health Study (NSPHS) that started in 2010 \cite{igl2010northern}. 
The approximately 1000 participants were initially mostly surveyed
about lifestyle \cite{igl2010northern} and follow-up studies
provided the type of data relevant for this paper, 
which are (1) the genetic data \cite{johansson2013identification},
(2) the concentration of certain proteins in the 
blood \cite{enroth2014strong,enroth2015effect}.

@@

The first type of data, the (final form of the) genetic data,
consists out of single nucleotide polymorphisms (SNPs).
SNPs consist out of a name, a position and a nucleotide. 
DNA (organized into chromosomes and present in every (nucleated) cell, 
see figure F3) consists out of many nucleotides, of which there are four types,
called adenosine, cytosine, guanine and thyrosine, all commonly abreviated
as A, C, G and T respectively.
One SNP example is, that 'rs12133641' is a SNP located at position 154,428,283,
where some people have a certain nucleotide. In this case, 
67 percent of the people within this study have an A,
and 33 percent have a G (also from \cite{ahsan2017relative}, Table S3).

@@

The field of genetic epidemiology looks -among others- for
correlations between genetic data and biological traits.
For example, figure F10 (from \cite{ahsan2017relative}) shows that
SNP rs12133641 is highly correlated (p value is 3.0 * 10e-73, 
961 individuals) with protein IL6RA. What this results does
not yet teach us, is how this correlation works, yet
figure F20 shows us the direction of the association:
the X axis shows the possible genetic make-ups (aka the genotype) of the individuals,
where 0 denotes the individuals with genotype
AA (the individual inherited one adenosine 
from his/here mother and one adenosine from his/her father), 
1.0 denotes AG (one A is inherited from one parent, 
where the G is inherited from the other parent) and 2.0 denotes GG.
From Figure F20 we can conclude that, on average,
the more guanines are inherited at that SNPs location,
the higher concentration of IL6RA can be found in a human's blood.

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{ahsan2017relative_table_2_sub.png}
  \caption{
    Figure F10 shows an example result of a genetic epideological research.
    It shows that the SNP named rs12133641 (located at position 154,428,283
    of chromosome 1) is highly correlated (p value is 3.0 * 10e-73, 
    961 individuals) to the concentration of the protein IL6RA, as measured
    in blood. The table is a simplified result from \cite{ahsan2017relative}.
  }
  \label{fig:ahsan2017relative_table_2_sub}
\end{figure}

@@

To quote a review on reproducibility \cite{peng2021reproducible}, 
'Reproducibility is typically thwarted by a 
lack of availability of the original data and computer code'.
In the case of genetic epidemiology, it is a given that the original
data cannot be published as this data is sensitive.

@@

The line of code is simplified (as the result of the T test is never stored),
but the code is correct:
\verb{t.test} is indeed the name of an R function to do a T-test
and it is reasonable to assume that this code, when found in the code
of an article, is assumed to be correct.

Here, however, we see a mismatch between the English description and the code:
by default, the R function \verb{t.test} does a \textbf{two}-tailed T-test.
A consequence of this fictional example is that the published p-values are
higher than needed, resulting in less significant findings, which
results in less conclusions drawn from a paper.

@@

The code that does the work in a computational experiment should
always be published together with a scholarly article.
For research to be reproducible, one ideally has access to
both the data used and the code.
In some fields, such as genetic epidemiology, the data is
sensitive, hence cannot be released,
yet there are methods being devised to run code on sensitive
data with assured privacy \cite{zhang2016review,azencott2018machine}.
